---
title: "Cross tables"
author: "Adrian Maldet"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: paged
vignette: >
  %\VignetteIndexEntry{cross tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Creating nicely styled **cross tables** in **R** can be tedios, even when you
use the `styledtable` package. Therefore, we implemented the following functions,
which make it very easy to create cross tables from [tidy data sets]:

-   `create_cross_table_header()`: 
    This function takes a `data.frame` and creates the styled header rows of a
    cross table. The cross table can have **multiple hirachy levels**.
-   `create_cross_table_body()`: 
    This function takes a `data.frame` and creates the styled body rows of a
    cross table. The cross table can have **multiple hirachy levels**.
-   `styled_table()`: 
    This function can be used to vertically concatenate the generated styled cross table header 
    with the generated styled cross table body.
  
## Creation of a tidy data set
We will need a [tidy data set] to illustrate the power of `styledtable` cross table creation.
Therefore, we create a `data.frame` named `students`, which holds some fictional test results
of students coming from **Germany** and **Austria** in subjects **Mathematics** and **Statistics** over several **years**. 
In the column `N` the number of cases and in the column `rel` the portion in percent is given.

```{r results = 'hide'}
library(dplyr)
students <- data.frame(
        country = rep(c("Germany", "Austria"), each = 16),
        year = c(rep("2010", 8), rep("2011", 8)),
        subject = rep(c(rep("Mathematics", 4), rep("Statistics", 4)), 2),
        gender = rep(c(rep("Male", 2), rep("Female", 2)), 4),
        result = rep(c("positive", "negative"), 16),
        N = sample(1:1000, 32)
    ) %>%
    group_by(country, year, subject, gender) %>%
    mutate(rel = 100 * N / sum(N)) %>%
    ungroup
```
```{r echo = F}
    knitr::kable(students)
```



[tidy data]: https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html
