---
title: "Getting started"
author: "Adrian Maldet"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

`styledTables` is an **R** package that makes it possible to export a
`data.frame` as beautifully styled **Excel** and **LaTeX** table.

> Produce tables, which you want to look at.

## Installation

```{r, eval = FALSE}
# Install development version from GitHub
devtools::install_github('R-package/styledTables', build_vignettes = TRUE)
```

## Create your first styled table

Use the `styled_table()` function to turn a `data.frame` or `matrix` into a
`StyledTable` class object.
Then apply all the [styling functions] you want to by using the [pipe operator] `%>%`
and return the `StyledTable` object at the end of your `Rmd` chunk so that `knitr` can
do the printing (the `print()` function of `knitr` is overloaded by `styledTables`).

```{r out.width = "8cm"}
library(styledTables)
library(dplyr)

# Create the styled table
st1 <- iris[sample(seq_len(nrow(iris)), 10),] %>%
    styled_table(keep_header = TRUE)

# Apply several styling functions
st1 <- st1 %>%
    set_border_position("all", row_id = 1) %>%
    set_bold(row_id = 1) %>%
    set_fill_color("#00FF00", col_id = 5, condition = X == "setosa")

# Return the styled table at the end of the Rmd-chunk in order to print it
st1
```

## Concat styled tables vertically
The function `styled_table()` allows you to put in multiple 
`data.frames`/`StyledTable` objects/`matrices`, which
will be concatenated vertically. This may be particularly useful for
adding a custimized table header to the table. With the functions `merge_cells()` 
and `merge_equal_cells()` multiple cells of a `StyledTable` object can be 
merged together.

```{r out.with = "6cm"}
library(styledTables)
library(dplyr)

# Create the styled table body
st_body <- iris[sample(seq_len(nrow(iris)), 10),] %>%
    styled_table %>%
    set_fill_color("#00FF00", col_id = 5, condition = X == "setosa")

# Create the styled table header
st_header <- c(
        "Sepal",
        "",
        "Petal",
        "",
        "Species",
        "Length",
        "Width",
        "Length",
        "Width",
        ""
    ) %>%
    matrix(nrow = 2, byrow = TRUE) %>%
    styled_table %>%
    set_border_position("all") %>%
    set_bold %>%
    merge_cells(col_id = 1:2, row_id = 1) %>%
    merge_cells(col_id = 3:4, row_id = 1) %>%
    merge_cells(col_id = 5, row_id = 1:2) %>%
    set_horizontal("center")

# Concatenate the styled table header and body
st2 <- styled_table(
    st_header,
    st_body
)

# Return the styled table at the end of the Rmd-chunk in order to print it
st2
```

## Export to Excel
For the creation of Excel tables the [xlsx package] is used. Create a `xlsx` 
workbook and use the `write_excel()` function to insert the `StyledTable` 
object in a worksheet.

```{r, eval = FALSE}
library(xlsx)

# open new xlsx workbook and create a worksheet
wb <- createWorkbook()
sheet <- createSheet(wb, "iris_table")

# insert the styled table in the worksheet
write_excel(sheet, st2)

# save the workbook
saveWorkbook(wb, "test.xlsx")
```
```{r, results = 'hide', echo = FALSE}
library(xlsx)

# open new xlsx workbook and create a worksheet
wb <- createWorkbook()
sheet <- createSheet(wb, "iris_table")

# insert the styled table in the worksheet
write_excel(sheet, st2)

# save the workbook
#saveWorkbook(wb, "../docs/ex2.xlsx")
```
[Resulting Excel file](../ex2.xlsx)

## Export to PDF or PNG
Use the function `write_pdf()` and `write_png()` to create `PDF` and `PNG`
images of the styled table.
```{r, eval = FALSE}
# Export to PDF
st2 %>%
    write_pdf("test.pdf")

# Export to PNG
st2 %>%
    write_png("test.png")
```

## Include in Rnw
The function `create_latex_table()` returns the `LaTeX` code of the `StyledTable` 
object. You can use this function to include your `StyledTable` inside of an `Rnw` file.
Therefore you must load several `LaTeX` packages in the `LaTeX` preamble
(before `\begin{document}`):
```{r, code=readLines("ex2.txt"), eval=FALSE}
```

```{r, include=FALSE, eval=FALSE}
system("cp ex2.txt ../docs/ex2.Rnw")
dir.create(tmp <- tempdir())
old_wd <- getwd()
# on.exit({
#     unlink(tmp, recursive = TRUE)
# })
system(
    paste(
        "cp", file.path(old_wd, "ex2.txt"), file.path(tmp, "ex2.Rnw"),
        "&& cd", tmp,
        paste0(
            "&& R --slave --no-init-file --no-save --no-restore -e \"setwd('",
            tmp, 
            "'); knitr::knit('ex2.Rnw')\""
        ),
        "&& pdflatex ex2.tex"
    )
)
system(paste("cp", file.path(tmp, "ex2.pdf"), "."))
system("cp ex2.pdf ../docs/")
[Rnw file](../ex2.Rnw) and [Resulting PDF file](../ex2.pdf)
```


[xlsx package]: https://cran.r-project.org/web/packages/xlsx/xlsx.pdf
[pipe operator]: https://style.tidyverse.org/pipes.html
